\mytitle

### 2020-04-22 paper.tex
evaluation.

To test and benchmark our models, we used 12 standard benchmark WFLO instances
from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement}. 
The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). We denote by $\mathcal{W}$ the set of 12 instances. 
The wind regime distribution presented in Figure~\ref{fig:prob_wind} was 
elicited from~\cite{Zhang2014}.
Additional parameters that were used to calculate 
the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.

@@

	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, color corresponds to free wind speed.}\label{fig:prob_wind}
\end{figure}

@@

We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in $kW/year$.}.
The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in~\cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different LS-based methods. For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742 kW/year$. Therefore, we let $b_w = 15742$. Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS\}$. Further, denote by $b_{w,t,a}$ denote the best solution by time $t$ of approach $a$ over instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by $$MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.$$ We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as $$MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).$$
We shall present this quantity ($MRE(t,a)$) for each of the approaches over time to demonstrate the speed of convergence of the proposed LS-based methods.

@@

In this part, we present the main findings of our empirical evaluation. 
We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW/hour) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. The runtime for DA-LS was always less than $10$ seconds.
We observe that...

Table~\ref{tab:results2} shows the same type of results after $3600$ seconds.

\begin{table}[t]

	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}           \\
		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
		&&40   & 21874.04 & 21643.56  & 21905.5 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
		&& 30  & \textbf{27450.99} & 27418.74 & 27442.9 & 27387.06 & 27247.5                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
		&&30   & \textbf{27975.55} & 27443.8  & 27752 & 27492.11  & 26367.2                      \\
		&&40   & 35604.45 & 35127.32 & \textbf{35665.6}   & 35065.85 & 33451.94 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method. * stands for optimal solution. \todo{@Jason: could you please check if my marking of * for optimal solutions is correct? I think that optimality was proven only for $m=20$ and single wind direction, right?} }\label{tab:results1}
\end{table}

\begin{table}[t]

@@

\vspace{0.5em}
\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
\end{table}

@@

	\subfloat[MRE after 50 seconds]{{\includegraphics[width=0.5\textwidth]{energy_50.png} }}%
	\quad
	\subfloat[MRE after 1 hour]{{\includegraphics[width=0.5\textwidth]{energy_3600.png} }}\\%\\
	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime.}%

	\label{fig:main_effects}%
\end{figure}

>>>>

To test and benchmark our models, we used 12 standard benchmark WFLO instances
from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement}. 
The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). 
For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}. We denote $\mathcal{W}$ the set of 12 instances. Note that instances with $36$ wind regimes are considered computationally harder~\cite{Zhang2014}.

Additional parameters that were used to calculate 
the velocities ($u_{id\infty}, u_{ijd}$) include: turbine height ($60$ meters),
ground roughness ($0.3$ meters), turbine rotor radius ($20$ meters), and a wind farm of size $2km \times 2km$.

@@

	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, and color corresponds to free wind speed for the different wind directions.}\label{fig:prob_wind}
\end{figure}

@@

We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in kW.}.
The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in~\cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different LS-based methods. For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$. Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS\}$. Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_{w,t,a}}{b_w}.\end{equation} We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $aMRE(t,a)$ can be computed as, \begin{equation}aMRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
We shall present $aMRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods. \todo{Jason: please check that this is consistent with how you calculated MRE in the experiment.}

@@

In this part, we present the main findings of our empirical evaluation. 
We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
The runtime for DA-LS was always less than $10$ seconds.

\begin{table}[t!]

	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & DA-LS  & QUBO-LS  & QP-LS  & QP-SS  & ILP-LS  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* & \textbf{11185.41}* \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.1  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.8}  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*  & \textbf{11404.8}*           \\
		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
		&&40   & 21874.04 & 21643.56  & 21905.5 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
		&& 30  & \textbf{27450.99} & 27418.74 & 27442.9 & 27387.06 & 27247.5                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
		&&30   & \textbf{27975.55} & 27443.8  & 27752 & 27492.11  & 26367.2                      \\
		&&40   & 35604.45 & 35127.32 & \textbf{35665.6}   & 35065.85 & 33451.94 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 100 seconds as function of benchmark and method. * stands for optimal solution. \todo{@Jason: could you please check if my marking of * for optimal solutions is correct? I think that optimality was proven only for $m=20$ and single wind direction, right?} }\label{tab:results1}
\end{table}

\begin{table}[t!]

@@

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 3600 seconds as function of benchmark and method. * stands for optimal solution.}\label{tab:results2}
\end{table}

We observe that the DA-based solution provides
comparable results to state-of-the-art exact methods. For the harder instances (of $36$ wind directions),
the three quadratic LS-based approaches (DA-LS, QUBO-LS, and QP-LS) dominate the others by finding the best (or equal) solutions in 11 out of 12 instances. For easier instances, 
the linear approach (ILP-LS) finds better solutions \todo{@Chris, Jason: do we have a good explanation for this?}. Table~\ref{tab:results2} shows the same type of results as in Table~\ref{tab:results1}, but now after $3600$ seconds. We see that the DA is still competitive with the other approaches (recall, its solution was found after $10$ seconds),
while QP-LS finds the best solutions in 10 out of the 12 scenarios. For the 6 hard scenarios (12 wind directions), our quadratic solutions
dominate all 6.

We now turn to present the $aMRE(a,t)$ over time: lower values indicate better performance. The different lines 
correspond to the 5 approaches. The horizontal axis represents time in seconds. Figure~\ref{fig:results1} 
presents the first $50$ seconds focusing on speed of performance, while Figure~\ref{fig:results2} corresponds
to the full run ($3600$ seconds).

@@

	\includegraphics[width=\textwidth]{energy_50.png}  
	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (first 50 seconds).}%

	\label{fig:results1}%
\end{figure}

	\includegraphics[width=\textwidth]{energy_3600.png}   \\%\\
	\caption{Mean Relative Energy (w.r.t. best feasible solution) vs. Runtime (3600 seconds).}%

	\label{fig:results2}%
\end{figure}
